---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Ben Hardin"
date: "1/31/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- dennys
laquinta <- laquinta
```

### Exercise 1

The Denny's dataset has 6 columns and 1643 rows. In this datafile, rows are individual Denny's. There are 6 variables (columns), representing each Denny's address, city, state, zip code, and longitude and latitude coordinates.

```{r dimensions-of-dennys}
#getting rows and columns of dennys
nrow(dennys)
ncol(dennys)
summary(dennys)
```

### Exercise 2

The Laquinta dataset has 6 columns and 909 rows. Again, the rows represent individual LaQuinta Inn locations. The 6 variables are the same ones we saw in the Denny's dataset, telling us the location of each Laquinta location. 

```{r dimensions-of-laquinta}
nrow(laquinta)
ncol(laquinta)
summary(laquinta)
```
### Exercise 3

From looking at the website, I can see that Denny's has no locations outside the US. LaQuinta on the other hand has several locations in other countries. These include countries like Canada, Chile, China, Columbia, New Zealand, Turkey, and the UAE.

### Exercise 4

We could also figure out whether these establishments have locations in other countries using just the datasets we have. Given the information we have, it seems like some options would be getting either the states or the zip codes for each location, and filtering for whether the state or zip code for each location is a place in the US or in a different country.

### Exercise 5

There are 0 Denny's locations left after filtering for states that are not in the US. This confirms what we saw earlier on the website, that there are no Denny's outside the US.

```{r states-of-dennys}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
### Exercise 6

The code chunk below makes a new variable for whether each Denny's is located in the US or another country. We now have a new dataset with 7 variables, one of which is our new country variable! Confirming our earlier findings, all 1643 observations in the dataset are Denny's located in the US.

```{r making-us-dennys-variable}
dennys <- dennys %>%
  mutate(country = if_else(state %in% states$abbreviation, "United States", "other"))

#did it work?
summary(dennys)

dennys %>%
  count(country)
```

### Exercise 7

Running similar code on the La Quinta dataset shows us that there are 14 La Quinta locations outside the US. Now, let's consult professor google to figure out which country each of these locations are in, and then plug those into the code for Ex. 8 to make a country variable.

```{r states-of-laquinta}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```

### Exercise 8

Great, now we have a country variable! We can see that, in this dataset, there are 10 La Quintas in Mexico, 2 in Canada, 1 in Honduras, and 1 in Columbia

```{r making-us-laquinta-variable}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% states$abbreviation  ~ "United States",
    state %in% c("ON", "BC")  ~ "Canada",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL")  ~ "Mexico",
    state == "ANT"  ~ "Colombia",
    state == "FM"  ~ "Honduras"
  ))

#did it work?
summary(laquinta)

laquinta %>%
  count(country)
```

### Exercise 9

Both Denny's and LaQuinta have a majority of their locations in California, Texas, and Florida. This isn't very surprising because these are 3 highly populated states. I also learned from professor google that the high concentration of locations in CA and TX can also be explained by the fact that Denny's and LaQuinta were founded in California and Texas, respectively.

```{r dennys-and-laquinta-by-state}
dennys %>%
  count(state) %>%
  arrange(desc(n))

laquinta %>%
  filter(country == "United States") %>%
  count(state) %>%
  arrange(desc(n))
```

### Exercise 10

Both Denny's and La Quinta have a large number of locations per square mile both in states where they have a lot of locations in general, like California, Texas, and Florida, as well as geographically small states like Rhode Island, Connecticut, and Maryland.

```{r locations-by-square-mile}
#dennys locations by square miles
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(locations_per_sqm = (n/area)) %>%
  arrange(desc(locations_per_sqm))

#laquinta locations by square miles
laquinta %>%
  filter(country == "United States") %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(locations_per_sqm = (n/area)) %>%
  arrange(desc(locations_per_sqm))
```

Now let's visualize the relationship between Denny's and La Quinta for the whole United States. Neat!

```{r visualizing}
#combining dennys and laquinta datasets
dennys <- dennys %>%
  mutate(establishment = "dennys")

laquinta <- laquinta %>%
  mutate(establishment = "laquinta")

dn_lq <- bind_rows(dennys, laquinta)

#visualizing relationship between dennys and laquinta
ggplot(dn_lq %>%
         filter(country == "United States"), mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment))+
  geom_point()
```

### Exercise 11

Now we can recreate the previous visualization showing only the Denny's and LaQuinta locations in North Carolina. In this case, there is some truth to Mitch Hedberg's joke, as most of the LaQuinta locations are pretty close to a Denny's, although there are a few stray LaQuintas that aren't near any Denny's, and there are many Denny's located in parts of NC that have no LaQuinta. 

```{r visualizing-NC}
ggplot(dn_lq %>%
         filter(state == "NC"), mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment))+
  geom_point(size = 3, alpha = 0.5)+
  labs(title = "Locations of Denny's and LaQuinta Inns in North Carolina")+
  theme(plot.title = element_text(hjust = 0.5))
```


### Exercise 12

Now let's create basically the same visualization, but this time for Texas. As we can see, Mitch Hedberg's joke doesn't hold quite as well here, because there are so many LaQuintas located in places where there is no Denny's. However, it does seem that, in places where there is a Denny's in Texas, there is likely to be a LaQuinta nearby. 

```{r visualizing-TX}
ggplot(dn_lq %>%
         filter(state == "TX"), mapping = aes(x = longitude,
                            y = latitude,
                            color = establishment))+
  geom_point(size = 3, alpha = 0.5)+
  labs(title = "Locations of Denny's and LaQuinta Inns in Texas")+
  theme(plot.title = element_text(hjust = 0.5))
```