---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Ben Hardin"
date: "1/31/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- dennys
laquinta <- laquinta
```

### Exercise 1

The Denny's dataset has 6 columns and 1643 rows. In this datafile, rows are individual Denny's. There are 6 variables (columns), representing each Denny's address, city, state, zip code, and longitude and latitude coordinates.

```{r dimensions-of-dennys}
#getting rows and columns of dennys
nrow(dennys)
ncol(dennys)
summary(dennys)
```

### Exercise 2

The Laquinta dataset has 6 columns and 909 rows. Again, the rows represent individual LaQuinta Inn locations. The 6 variables are the same ones we saw in the Denny's dataset, telling us the location of each Laquinta location. 

```{r dimensions-of-laquinta}
nrow(laquinta)
ncol(laquinta)
summary(laquinta)
```
### Exercise 3

...

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

 
Add exercise headings as needed.

